---
title: 'BT Question Set P1-T2-20-19: Regression diagnostics (1st set)'
author: David Harper
date: '2020-09-21'
slug: bt-question-set-p1-t2-20-19-regression-diagnostics-1st-set
categories:
  - R
tags:
  - regression
subtitle: ''
summary: ''
authors: []
lastmod: '2020-09-21T12:37:21-07:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

## P1.T2.20.19. Regression diagnostics: omitted variables, heteroskedasticity, and multicollinearity

### Question 1: Fama-french 2-factor <- omitted variable

20.19.1. Jane manages a market-neutral equity fund for her investment management firm. The fund's market-neutral style implies (we will assume) that the fund's beta with respect to the market's excess return is zero. However, the fund does seek exposure to other factors. The size factor captures the excess return of small-capitalization stocks (SMB = "small minus big"). Jane tests her portfolio's exposure to the size factor by regressing the portfolio's excess return against the size factor returns. Her regression takes the form PORTFOLIO(i) = α + β1×SMB(i) + ε(i). The results of this single-variable (aka, simple) regression are displayed below.

... univariate regression output ...

In this simple regression, we can observe that SMB's coefficient is 0.6771 and significant. Jane is concerned that this simple regression might suffer from omitted variable bias. Specifically, she thinks the value factor has been omitted. The value factor captures the excess returns of value stocks (HML = "high book-to-market minus low book-to-market")'. She confirms that the omitted variable, HML, is associated with her response variable. Further, the omitted variable, HML, is correlated to SMB. The correlation between HML and SMB is 0.30. Jane runs a multivariate regression with both explanatory variables, SMB and HML; in this regression, HML's beta coefficient is 0.7240 such that the new term is β2×HML(i) = 0.7240×HML(i). Which of the following is nearest to the revised SMB coefficient; i.e., what is the revised β1?

```{r warning=FALSE, message=FALSE}

library(tidyverse)
library(broom)
library(gt)

intercept <- .03
intercept_sig <- .01

# x1 is market factor
x1_mu <- .05
x1_sig <- .01
x1_beta <- 0.5

# x2 is size factor
x2_mu <- .04
x2_sig <- .01
x2_beta <- 0.7

#x3 is value factor
x3_mu <- .03
x3_sig <- .01
x3_beta <- -0.5

noise_mu <- 0
noise_sig <- 0 # low value gets low p-value b/c low noise

size <- 120
set.seed(52)

results <- tibble(
  x0 = rnorm(size, intercept, intercept_sig),
  x1 = rnorm(size, x1_mu, x1_sig),
  x2 = rnorm(size, x2_mu, x2_sig),
  x2_r = 0.3 * x1 + sqrt(1 - 0.3^2)* x2,
  x3 = rnorm(size, x3_mu, x3_sig),
  x1_b = rep(x1_beta, size),
  x2_b = rep(x2_beta, size),
  x3_b = rep(x3_beta, size),
  noise = rnorm(size, 0, noise_sig)
)

results1 <- results %>% mutate(
  y1 = x0 + x1_b * x1 + noise, # market factor only
  y2 = x0 + x1_b * x1 + x2_b * x2_r + noise, # market and size
  y3 = x0 + x1_b * x1 + x2_b * x2 + x3_b * x3 + noise # all three
)

model_y1 <- lm(y2 ~ x1, data = results1) # market only
model_y2 <- lm(y2 ~ x1 + x2_r, data = results1) # market and size
model_y3 <- lm(y3 ~ x1 + x2 + x3, data = results1)

cor_x1_x2 <- cor(results1$x1, results1$x2_r)
cov_x1_x2 <- cov(results1$x1, results1$x2_r)
var_x1 <- var(results1$x1)
var_x2 <- var(results1$x2_r)
beta_x1 <- cov_x1_x2/var_x1

model_omit <- lm(x2_r ~ x1, data = results1)

# model_y1 suffers omitted variable bias: x1 = 0.6771 as given in problem
summary(model_y1)


# And note that "true" X1 = 0.463 per the solution's 0.460
# Problem provides rounded 0.7240 as HML coefficient
summary(model_y2)


model_tidy <- tidy(model_y1)
model_tidy[2,1] <- "SMB"

gt_table_model <- gt(model_tidy)

gt_table_model <- 
  gt_table_model %>% 
  tab_options(
    table.font.size = 14
  ) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body()
  ) %>% 
  tab_header(
    title = "Market-neutral portfolio excess returns regressed against SMB",
    subtitle = md("(but HML is an ***omitted variable***)")
  #) %>% tab_source_note(
  #   source_note = md("the source is ... FRED")
  ) %>% cols_label(
    term = "Coefficient",
    estimate = "Estimate",
    std.error = "Std Error",
    statistic = "t-stat",
    p.value = "p value"
  ) %>% fmt_number(
    columns = vars(estimate, std.error, statistic, p.value),
    decimals = 4
  ) %>% fmt_scientific(
    columns = vars(statistic, p.value),
  ) %>% tab_options(
    heading.title.font.size = 14,
    heading.subtitle.font.size = 14
  )

gt_table_model

```

## Question 2: House Prices - Heteroskedasticity

20.19.2. Josh regressed house prices (as the response or dependent variable) against two explanatory variables: square footage (SQFEET) and the number of rooms in the house (ROOMS). The dependent variable, PRICE, is expressed in thousands of dollars ($000); e.g., the average PRICE is $XXX.000 because the average house price in the sample of 150 houses is $XXX.000. The units of SQFEET are unadjusted units; e.g., the average SQFEET in the sample is X,XXX ft^2. The variable ROOMS is equal to the sum of the number of bedrooms and bathrooms; because much of the sample is 2- and 3-bedroom houses with 2 baths, the average of ROOM is X.XX. Josh's regression results are displayed below.

Josh is concerned that the data might not be homoscedastic. He decides to conduct a White test for heteroskedasticity. In this test, he regresses the squared residuals against each of the explanatory variables and the cross-product of the explanatory variables (including the product of each variable with itself). The results of this regression are displayed below.

```{r warning=FALSE, message=FALSE}

# library(tidyverse)
# library(broom)
# library(gt)

intercept <- 40
intercept_sig <- .01

x1_mu <- 1000
x1_sig <- 400
x1_beta <- 0.3

x2_mu <- 4.5
x2_sig <- 2
x2_beta <- 20.0

noise_mu <- 0
noise_sig <- 500 # low value gets low p-value b/c low noise

size <- 150
set.seed(65)

rho_noise_x1 <- 0.7

ft2_start = 800
ft2_end = 3000

results <- tibble(
  x0_sn = rnorm(size),
  x1_sn = rnorm(size),
  x2_sn = rnorm(size),
  e = rnorm(size),
  # e_corr = rho_noise_x1 * x1_sn + sqrt(1 - rho_noise_x1^2) * noise_sn,
  # 
  x0 = intercept + x0_sn * intercept_sig,
  # x1 = x1_mu + x1_sn * x1_sig, 
  x1 = seq(ft2_start, ft2_end - 1, by = (ft2_end - ft2_start)/size),
  x2 = x2_mu + x2_sn * x2_sig,
  # 
  # here the heteroskadisticity is deliberately introduced:
  # the error term's sigma is increasing with SQFEET, x1
  e_sigma = ((x1 - ft2_start)/100)^2, 
  # noise = noise_sn_corr * noise_sig, 
  #  
  x1_b = rep(x1_beta, size),
  x2_b = rep(x2_beta, size)
)

results1 <- results %>% mutate(
  y = x0 + (x1_b * x1) + (x2_b * x2) + (e * e_sigma)
)

# cor(results1$e * results1$e_sigma, results1$x1)

model_house <- lm(y ~ x1 + x2, data = results1)
summary(model_house)

model_tidy_house <- tidy(model_house)
model_tidy_house[2,1] <- "SQFEET"
model_tidy_house[3,1] <- "ROOMS"

gt_table_model_house <- gt(model_tidy_house)

gt_table_model_house <- 
  gt_table_model_house %>% 
  tab_options(
    table.font.size = 12
  ) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body()
  ) %>% 
  tab_header(
    title = "House Price regressed against ft^2 (SQFEET) + ROOMS(#)",
    subtitle = md("House Price in Thousands **($000)** of dollars")
  ) %>% tab_source_note(
    source_note = md("Residual standard error: 211.1 on 147 degrees of freedom")
  ) %>% tab_source_note(
    source_note = md("Multiple R-squared:  0.5548,	Adjusted R-squared:  0.5488")
  ) %>% tab_source_note(
    source_note = md("F-statistic: 91.61 on 2 and 147 DF,  p-value: < 2.2e-16")  
  ) %>% cols_label(
    term = "Coefficient",
    estimate = "Estimate",
    std.error = "Std Error",
    statistic = "t-stat",
    p.value = "p value"
  ) %>% fmt_number(
    columns = vars(estimate, std.error, statistic, p.value),
    decimals = 3
  ) %>% fmt_scientific(
    columns = vars(p.value),
  ) %>% tab_options(
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    source_notes.font.size = 12
  )

gt_table_model_house

mean(results1$y) # price
mean(results1$x0) # intercept
mean(results1$x1) # sqfeet
mean(results1$x2) # rooms

ressq <- resid(model_house)^2
results_v2 <- cbind(results1, ressq)
white_test <- lm(ressq ~ x1 * x2 + I(x1^2) + I(x2^2), data = results_v2)
summary(white_test)

white_tidy <- tidy(white_test)
white_tidy[2,1] <- "SQFEET"
white_tidy[3,1] <- "ROOMS"
white_tidy[4,1] <- "SQFEET^2"
white_tidy[5,1] <- "ROOMS^2"
white_tidy[6,1] <- "SQFEET*ROOMS"

gt_table_white <- gt(white_tidy)

gt_table_white <- 
  gt_table_white %>% 
  tab_options(
    table.font.size = 12,
  ) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body()
  ) %>% 
  tab_header(
    title = "RESIDUAL^2 regressed against SQFEET + ROOMS + SQFEET^2 + ROOMS^2 + ROOMS*SQFEET",
    subtitle = md("White's Test for Heteroskedasticity")
  ) %>% tab_source_note(
    source_note = md("Residual standard error: 76920 on 144 degrees of freedom")
  ) %>% tab_source_note(
    source_note = md("Multiple R-squared: 0.3381,	Adjusted R-squared: 0.3152, F-statistic: 14.71 on 5 and 144 DF, p-value: 1.201e-11")
  ) %>% cols_label(
    term = "Coefficient",
    estimate = "Estimate",
    std.error = "Std Error",
    statistic = "t-stat",
    p.value = "p value"
  ) %>% fmt_number(
    columns = vars(estimate, std.error, statistic, p.value),
    decimals = 3
  ) %>% fmt_scientific(
     columns = vars(estimate, std.error),
  ) %>% tab_options(
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    source_notes.font.size = 12
  )

gt_table_white

# and the heteroskadisticity is visually VERY obvious!
results_v2 %>% ggplot(aes(x = x1, y = y)) +
  geom_point()

```

## Question 3: Multicollinearity

20.19.3. Emily works for an insurance company and she has regressed medical costs (aka, the response or dependent variable) for a sample of patients against three independent variables: AGE, BMI, and CHARITY. The sample's average age is X.X years. Body mass index (BMI) is mass divided by height squared and the sample's average BMI is X.Y. CHARITY is the dollar amount of charitable spending in the last year; the sample average is $X.Y donated to charity in the last year. Emily's regression results are displayed below.

```{r warning=FALSE, message=FALSE}

library(tidyverse)
library(broom)
library(gt)

intercept <- 150
intercept_sig <- 40

# age
x1_mu <- 38
x1_sig <- 7
x1_beta <- 50

# bmi
x2_mu <- 22
x2_sig <- 4
x2_beta <- 100

# spend
x3_mu <- 500
x3_sig <- 250
x3_beta <- -0.4

rho_x1_x2 = 0.3
rho_x1_x3 = 0.97

noise_mu <- 0
noise_sig <- 300 # low value gets low p-value b/c low noise

size <- 43
set.seed(12)

results <- tibble(
  x0_sn <- rnorm(size),
  x1_sn <- rnorm(size),
  x2_sn <- rnorm(size),
  x3_sn <- rnorm(size),
  #
  x0 = intercept + intercept_sig * x0_sn,
  #
  x1 = x1_mu + x1_sig * x1_sn,
  x2_sn_corr = rho_x1_x2*x1_sn + sqrt(1 - rho_x1_x2^2)*x2_sn,
  x2 = x2_mu + x2_sig * x2_sn_corr,
  x3_sn_corr = rho_x1_x3*x1_sn + sqrt(1 - rho_x1_x3^2)*x3_sn,
  x3 = x3_mu + x3_sig * x3_sn_corr,
  #
  x1_b = rep(x1_beta, size),
  x2_b = rep(x2_beta, size),
  x3_b = rep(x3_beta, size),
  noise = rnorm(size, 0, noise_sig)
)


results1 <- results %>% mutate(
  y = x0 + (x1_b * x1) + (x2_b * x2) + (x3_b * x3) + noise
)

cor(results1$x1, results1$x3)

model <- lm(y ~ x1 + x2 + x3, data = results1)

model_x1 <- lm(x1 ~ x2 + x3, data = results1)
summary(model_x1)
rsq_x1 <- summary(model_x1)$r.squared

model_x2 <- lm(x2 ~ x1 + x3, data = results1)
summary(model_x2)
rsq_x2 <- summary(model_x2)$r.squared

model_x3 <- lm(x3 ~ x1 + x2, data = results1)
summary(model_x3)
rsq_x3 <- summary(model_x3)$r.squared

var_if_x1 <- 1/(1 - rsq_x1 )
var_if_x2 <- 1/(1 - rsq_x2)
var_if_x3 <- 1/(1 - rsq_x3)

multi_table <- tibble(
  regression = c(1,2,3),
  response = c("AGE", "BMI", "CHARITY"),
  explanatory = c("BMI + CHARITY", "AGE + CHARITY", "AGE + BMI"),
  Rsquared = c(rsq_x1, rsq_x2, rsq_x3)
)


model_tidy <- tidy(model)
model_tidy[2,1] <- "AGE"
model_tidy[3,1] <- "BMI"
model_tidy[4,1] <- "CHARITY"

gt_table_model <- gt(model_tidy)

gt_table_model <- 
  gt_table_model %>% 
  tab_options(
    table.font.size = 12
  ) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body()
  ) %>% 
  tab_header(
    title = "Medical COST regressed against AGE + BMI + CHARITY($)",
    subtitle = md("Simulated data")
  ) %>% tab_source_note(
    source_note = md("Residual standard error: 325.1 on 39 degrees of freedom")
  ) %>% tab_source_note(
    source_note = md("Multiple R-squared: 0.6961,	Adjusted R-squared: 0.6727, F-statistic: 29.77 on 3 and 39 DF, p-value: 3.514e-10")
  ) %>% cols_label(
    term = "Coefficient",
    estimate = "Estimate",
    std.error = "Std Error",
    statistic = "t-stat",
    p.value = "p value"
  ) %>% fmt_number(
    columns = vars(estimate, std.error, statistic, p.value),
    decimals = 2
  ) %>% fmt_scientific(
    columns = vars(p.value),
  ) %>% tab_options(
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
     source_notes.font.size = 12
  )

##

gt_multi_table <- gt(multi_table)

gt_multi_table <- 
  gt_multi_table %>% 
  tab_options(
    table.font.size = 12
  ) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body()
  ) %>% 
  tab_header(
    title = "Each response variable regressed against the others",
    subtitle = md("Testing for multicollinearity")
  ) %>% tab_source_note(
    source_note = md("Note: According to GARP, the standard test of")
  ) %>% tab_source_note(
    source_note = md("multicollinearity is the variance inflation factor (VIF)")
  ) %>% cols_label(
    regression = "Regression",
    response = "Response",
    explanatory = "Explanatory",
    Rsquared = "R-squared",
  ) %>% cols_align(
    align = "center",
    columns = vars(regression)
  ) %>% fmt_number(
    columns = vars(Rsquared),
    decimals = 3
  # ) %>% fmt_scientific(
    # columns = vars(p.value),
  ) %>% tab_options(
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
     source_notes.font.size = 12
  )

summary(model)
summary(multi_table)

gt_table_model
gt_multi_table

mean(results1$y) # cost
mean(results1$x0) # intercept
mean(results1$x1) # age
mean(results1$x2) # bmi
mean(results1$x3) # spend

# let's see the VIFs
var_if_x1
var_if_x2
var_if_x3

```


